---
title: "Explore and Mine Data"
author: "Prathamesh jagdale"
date: "Spring2024"
output: html_notebook
---

```{r Connecting to the database}
library(DBI)
library(kableExtra)
library(ggplot2)
library(dplyr)
library(knitr)
library(DT)


# Establish MySQL database connection
con <- dbConnect(
  RMySQL::MySQL(),
  dbname = "practicum2",
  host = "database-2.cr84qqy6ksk7.us-east-2.rds.amazonaws.com", 
  port = 3306, 
  user = "admin",
  password = "practicum2"
)

# Checking if the connection has been successfully established
if (dbIsValid(con)) {
  cat("Connection to MySQL database successful!\n")
} else {
  cat("Failed to connect to the MySQL database.\n")
}
```
###Analytical Query I: Top five sales reps with the most sales broken down by year. So, for each year for which there is data, the top five reps for that year. Think about how you would best present this information so it is easily usable by an analyst or manager
```{r Analytical Query 1}
query1 <- "
WITH ranked_reps AS (
    SELECT 
        year,
        repName,
        SUM(totalSold) AS total_sales,
        ROW_NUMBER() OVER (PARTITION BY year ORDER BY SUM(totalSold) DESC) AS sales_rank
    FROM 
        rep_facts 
    GROUP BY 
        year, repName
)
SELECT 
    year,
    repName,
    total_sales
FROM 
    ranked_reps
WHERE 
    sales_rank <= 5
ORDER BY 
    year, total_sales DESC;

"
result1 <- dbGetQuery(con, query1)
# Formatting the data for presentation
 datatable(result1, 
          options = list(
            stripeClasses = c('odd', 'even'),  # Apply alternating row colors
            paging = TRUE,                      # Enable pagination if needed
            searching = TRUE                    # Enable search functionality
          )
)


```

###Analytical Query II: Total sold per product per quarter. Think about how to best display this information.
```{r Analytical Query II}
AQ2 <- "
SELECT 
    productName,
    quarter,
    SUM(totalUnitsSold) AS totalUnitsSold
FROM product_facts 
GROUP BY productName, quarter
ORDER BY
    productName,
    quarter;
"
# Execute the query and store the result
resultII <- dbGetQuery(con, AQ2)

# Arrange the result by productName, and quarter
resultII <- arrange(resultII, productName, quarter)

datatable(resultII, 
          options = list(
            paging = TRUE,         # Enable pagination
            searching = TRUE,      # Enable search functionality
            ordering = TRUE       # Enable column sorting
          ),
          caption = "Average Sales per Sales Rep Over the Years"
)



```


### Analytical Query III: Number of units sold per product per region. Show the result in a line graph visualization
```{r Analytical_Query_III, warning=FALSE}

# Execute the query and store the result
AQ3 <- "
SELECT 
    productName,
    territory,
    SUM(totalUnitsSold) AS totalUnitsSold
FROM product_facts 
GROUP BY productName, territory
"
resultIII <- dbGetQuery(con, AQ3)

# Arrange the result 
resultIII <- arrange(resultIII, productName, territory)

# Formatting the data for presentation
result <- resultIII %>%
  kable("html") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

result


library(plotly)

# Assuming 'result3' is your data frame containing the query result

# Create the line plot
line_plot <- plot_ly(resultIII, x = ~territory, y = ~totalUnitsSold, color = ~productName, type = 'scatter', mode = 'lines+markers') %>%
  layout(title = "Total Units Sold by Territory",
         xaxis = list(title = "Territory"),
         yaxis = list(title = "Total Units Sold"),
         legend = list(orientation = "h"))

# Print the line plot
line_plot

```



###Analytical Query IV: Average sales per sales rep over the years. Show the result in a line graph visualization.
```{r Analytical Query IV}
library(DT)
# Execute the SQL query and fetch the result
AQ4 <- "
SELECT 
    rf.repName,
    rf.year,
    AVG(rf.totalSold) AS avg_sales
FROM 
    rep_facts rf
JOIN 
    product_facts pf ON rf.productID = pf.productID
GROUP BY 
    rf.repName, rf.year;
"
resultIV <- dbGetQuery(con, AQ4)

# Display the result in table format
datatable(resultIV, 
          options = list(
            paging = TRUE,         # Enable pagination
            searching = TRUE,      # Enable search functionality
            ordering = TRUE       # Enable column sorting
          ),
          caption = "Average Sales per Sales Rep Over the Years"
)
library(ggplot2)

# Plot the line graph
ggplot(resultIV, aes(x = year, y = avg_sales, group = repName, color = repName)) +
  geom_line() +
  geom_point() +
  labs(title = "Average Sales per Sales Rep Over the Years",
       x = "Year",
       y = "Average Sales") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_color_viridis_d()

dbDisconnect(con)

```
## Conclusion

This report provides valuable insights into our sales performance, highlighting trends and patterns that can inform strategic decision-making. By understanding our sales data more comprehensively, we can identify opportunities for growth and optimize our sales strategies.





